---
title: "820 project"
output:
  html_document: default
  word_document: default
  pdf_document: default
editor_options: 
  chunk_output_type: inline
---
setwd("~/")

Reading the data to be assesed, as well as cleaning and assigning proper titles to columns


```{r}
RV_data <- read.csv("Runs value data.csv")

Year <- RV_data$ï..Year
HR_Run_Value <- RV_data$Home.Run.Runs.Value
HR_Percentage <- RV_data$Home.Run.Percentage
Single_Run_Value <- RV_data$Single.Runs.Value
Single_Percentage <- RV_data$Single.Percentage
SB_Run_Value <- RV_data$Stolen.Base.Runs.Value
SB_Percentage <- RV_data$Stolen.Base.Percentage
SB_APG <- RV_data$Stolen.Base.Attempts.Per.Game
Strikeout_Percentage <- RV_data$Stikeout.Percentage

```

Initial Assessment of data, to analyze univariate data of each attribute being assessed.


```{r}
library("stats")
library("psych")

summary(Year)

summary(HR_Run_Value)
describe(HR_Run_Value)

summary(HR_Percentage)
describe(HR_Percentage)

summary(Single_Run_Value)
describe(Single_Run_Value)

summary(Single_Percentage)
describe(Single_Percentage)

summary(SB_Run_Value)
describe(SB_Run_Value)

summary(SB_Percentage)
describe(SB_Percentage)

summary(SB_APG)
describe(SB_APG)

summary(Strikeout_Percentage)
describe(Strikeout_Percentage)

plot(RV_data)
```

Testing the correlation between the variables

```{r}
#Overall Correlation of entire data set
cor(RV_data)

#Correlation between the run value of a home run and the percentage of plate appearances that result in a home run. 
cor.test(HR_Run_Value, HR_Percentage)

#Correlation between the run value of a single and the percentage of hits that are singles.
cor.test(Single_Run_Value, Single_Percentage)

#Correlation between the run value of a stolen base and the percentage of stolen base attempts that are successful. 
cor.test(SB_Run_Value, SB_Percentage)

```

Assessing the change of each variable over time. 


```{r}
#Run Value of a Home run over time
plot(Year, HR_Run_Value, ylab = "HR Run Value", main= "Year vs HR Run Value") + abline(lm(HR_Run_Value~Year, data=RV_data))
summary(lm(Year~HR_Run_Value))

#Home Run Percentage over time
plot(Year, HR_Percentage, ylab = "HR Run Percentage", main= "Year vs HR Percentage") + abline(lm(HR_Percentage~Year, data=RV_data))
summary(lm(Year~HR_Percentage))

#Run Value of a single over time
plot(Year, Single_Run_Value, ylab = "Single Run Value", main= "Year vs Single Run Value") + abline(lm(Single_Run_Value~Year, data=RV_data))
summary(lm(Year~Single_Run_Value))

#Single percentage over time
plot(Year, Single_Percentage, ylab = "Single Percentage", main= "Year vs Single Percentage") + abline(lm(Single_Percentage~Year, data=RV_data))
summary(lm(Year~Single_Percentage))

#Run Value of a stolen base over time
plot(Year, SB_Run_Value, ylab = "Stolen Base Run Value", main= "Year vs Stolen Base Run Value") + abline(lm(SB_Run_Value~Year, data=RV_data))
summary(lm(Year~SB_Run_Value))

#Stolen Base percentage over time
plot(Year, SB_Percentage, ylab = "Stolen Base Percentage", main= "Year vs Stolen Base Percentage") + abline(lm(SB_Percentage~Year, data=RV_data))
summary(lm(Year~SB_Percentage))

#Stolen Base attempts per game over time
plot(Year, SB_APG, ylab = "SB Attempts Per Game", main= "Year vs SB Attempts Per Game") + abline(lm(SB_APG~Year, data=RV_data))
summary(lm(Year~SB_APG))

#Strikeout Percentage over time
plot(Year, Strikeout_Percentage, ylab = "Strikeout Percentage", main= "Year vs Strikeout Percentage") + abline(lm(Strikeout_Percentage~Year, data=RV_data))
summary(lm(Year~Strikeout_Percentage))

```


Predictive Analysis

```{r}
#building linear regression model
library(caTools)

sample.split(HR_Run_Value, SplitRatio = 0.70) -> split_tag
subset(RV_data, split_tag==T) -> train
subset(RV_data, split_tag==F) -> test

lm(HR_Run_Value~Year, data = train) -> model1
predict(model1, newdata = test) -> predicted_values
head(predicted_values)

cbind(Actual= test$Home.Run.Runs.Value, Predicted= predicted_values) -> final_data
as.data.frame(final_data)-> final_data

final_data$Actual - final_data$Predicted -> error
cbind(final_data,error) ->final_data
head(final_data)
rmse <- sqrt(mean(final_data$error)^2)
rmse
```

```{r}
# multi linear regression model
library(caTools)

sample.split(HR_Run_Value, SplitRatio = 0.75) -> split_model
subset(RV_data, split_model==T) -> train2
subset(RV_data, split_model==F) -> test2

lm(HR_Run_Value~Year+HR_Percentage+Single_Run_Value+Single_Percentage+SB_Run_Value+SB_Percentage+SB_APG+Strikeout_Percentage, data = train2) -> mod_multi_linear

predict(mod_multi_linear, newdata = test2) -> predicted_multi_linear
head(predicted_multi_linear)

cbind(Actual= test2$Home.Run.Runs.Value, Predicted = predicted_multi_linear) -> final_data2
as.data.frame(final_data2) ->final_data2
head(final_data2)

final_data2$Actual - final_data2$Predicted -> error2
head(error)

cbind(final_data2,error2) ->final_data2
head(final_data2)

rmse_ml<- sqrt(mean((final_data2$error2)^2))
rmse_ml
```
Lower model presented above is the first linear regression model when assesing the run value of a home run. 


